---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

```

# wizard

A set of convenience functions for everyday programming and for visualization. Work in progress. Currently just a bunch of functions that I use very often.

## qbarplot

The `qbarplot()` function allows for quickly generating ggplot2 barplots from raw dataframes:

```{r example, eval = FALSE}

qbarplot(diamonds, color, cut)

```


## qbeeswarm

The `qbeeswarm()` function allows for quickly generating ggplot2 beeswarm plots with overlaid boxplots. Depends on the ggbeeswarm package.

```{r example2, eval = FALSE}

set.seed(200)
df <- data.frame(A = c(rep("high", 50), rep("lo", 50)),
                 B = round(runif(n = 100)*100))
qbeeswarm(df, x = A, y = B)

```

## pretty_df

The `pretty_df()` function formats the output of dataframes in a way that there are two digits after the decimal separator. Note that the "prettified" version will often not be printed in the R output but it will appear e.g. if you export the data to a CSV file.

```{r example3, eval = FALSE}

df <- data.frame(a = c("a", "b", "c"),
                 b = c(0.89469435394539469569, 123684.2683523825, 0.00000005))

p_df <- pretty_df(df)
write.csv(p_df, "prettydataframe.csv")

```


## format_age

This might be helpful for researchers in the field of language acquisition, where we often work with age values formatted like "years;months.days", e.g. "2;3.4". R will usually parse these as character values, and if there are no trailing zeros (as in "02;03.04"), they will be sorted incorrectly (so that a child at 02;10.17 is suddenly younger than a child at 2;1.11). This function circumvents this problem by adding trailing zeros. If you prefer the formatting without trailing zeros, use `factors = TRUE` and the factor levels of the original age column will be ordered correctly.

```{r example4, eval=FALSE}

# for vectors
x <- c("1;3.4", "1;2.5", "1;4.3")
format_age(x)

# for dataframes:
df <- c(age = c("1;3.4", "1;2.5", "1;4.3"),
        utterance = "bla", "bla", "bla")

format_age(df,
           col = "age", # name of the age column
           month_separator = ";", # sign separating month from year
           day_separator = ".", # sign separating day from month
           factors = T # reorder factor levels of original column (default is FALSE)
           )

# if the column name of the dataframe is "age" or "Age"
# (or is the only column containing the string "age" or "Age"),
# you don't have to specify the column name:
df <- c(age = c("1;3.4", "1;2.5", "1;4.3"),
        utterance = "bla", "bla", "bla")

format_age(df)


````


## load_packages

The `load_packages()` function checks whether the packages specified as its arguments are installed and installs them if they are not. If `require = TRUE` (the default), they are also loaded in the process.


## To be continued

More functions following soon!
